\documentclass[defaultstyle,12pt]{thesis}

\usepackage{amssymb}		% to get all AMS symbols
\usepackage{graphicx}		% to insert figures

%%%%%%%%%%%%   All the preamble material:   %%%%%%%%%%%%

\title{Using Rule Induction to Elucidate Co-Occurrence Patterns in Microbial
Data}

\author{K.~Kumar}{Thurimella}

\degree{Bachelors of Science}		%  #1 {long descr.}
	{B.S., Applied Mathematics}		%  #2 {short descr.}

\dept{Department of}			%  #1 {designation}
	{Applied Mathematics}		%  #2 {name}

\advisor{Associate Professor}				%  #1 {title}
	{Rob Knight}			%  #2 {name}

\reader{Professor~Michael Mozer}		%  2nd person to sign thesis
\readerThree{Senior~Instructor~Anne Dougherty}		%  3rd person to sign thesis

\abstract{  \OnePageChapter	% because it is very short
    Several studies have addressed whether the presence or absence of certain bacteria are
    linked with a particular phenotype. However, it is plausible that
    the causitive agent (or the consequence) of a given phenotype is not
    a single microbe, but groups of them. Rule Induction is a commonly used
    machine learning tool to infer structure within observational
    data and build rules to represent respective structures. In this thesis I
    introduce the application of a method, Rule Induction, to infer co-occurrence patterns in microbial data.
    \\ \indent First, I benchmark the methods within Rule Induction and
    understand how rules are generated with regards to several parameters such
    as table density, support and confidence. I then subsample data over
    multiple iterations to understand the robustness of the rules being
    presented and preserved over each sampling. 
    \\ \indent Next, I provide insight into different biological variables and
    examine their effect on rules produced. I compare 16s rRNA region,
    specficially V1-3 and V3-5 regions. I compare different sequencing
    technology, specifically 454 and Illumina. I finally compare time,
    specifically looking over 400 days. Within all these comparisons I aim to
    understand the differences, but more importantly what is conserved within
    these variables by the means of rules generated.
    
    \\ \indent Finally, I explore Rule Induction on two microbial datasets and
    see how strong the rules are in comparison to already known associations.
    The first dataset I interpret regards a correlation between HIV and the Gut
    Microbiome. The second dataset distinguishes the Gut Microbiome over varying
    georgraphical locations. I link each of these rules produced from each
    dataset with taxonomic information and consildate those rules to give rise
    the underlying structure within the biological data. }

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	To my family and friends.
	}

\acknowledgements{
	\indent First and foremost I would like to thank Rob Knight
	for letting me be a part of his lab. He has been a fantastic mentor
	and someone who I look up to very much. His passion for research is contageous
	and I have learned so much in my time here, and am very fascinated and
	intruiged by the research being done in the microbiome field in addition 
	to this lab. I would like to thank Mike Mozer for being a great
	mentor as well, by giving me great advice specific to this project
	as well as general life advice. My final committee member, Anne Dougherty,
	has been nothing short of a phenomenal advisor. It is after talking to her
	my sophomore year that I switched my major to Applied Mathematics. She has
	consistently provided great support and I can go to her about anything like 
	another friend.
	\\ \indent I would like to thank Jose Clemente for his ideas, support and
	fabulous mentoring. I have learned a lot from Jose with regards to his perspective 
	on being a researcher. Will van Trueren has been nothing but a great friend and
	peer mentor in the lab all the while providing great insights to this research.
	I want to finally thank other members of this lab including Yoshiki Baeza for
	his help understanding cluster computing, Cathy Lozupone for her HIV data and
	insight within co-occurrence and many others who were always there for
	support. I want to acknowledge the Gautam Dantas Lab at
	Washington University in St. Louis, specifically Kevin Forsberg and Mitch Pesesky. This past summer
	opened my eyes to exciting avenues of research and that experience provided the
	direction of research that I continue to this day.
	\\ \indent Without my close friends I wouldn't be where I
	am at today. Thanks to my roommates David Gillis and Thomas Lynn for helping
	me out this year. Oriel Eisner for always being interested in (or putting
	with) our late night talks, mostly regarding science. Many thanks Sathish Subramanian for being a
	fantastic role model and providing help and support whenever I needed it, 
	as well as late night life talks. I hope to be half the MD/PhD student he is,
	one day. Will Timbers has always been incredibly fun to talk with and has
	pushed me to pursue my passions. Andrew Fleming has shared a same passion for
	science with me in high school and I appreciate all of his support over the years. Without Myke Samuels I don't know where I would be, and it was because of his help that I
	found my passion and interests and for that I am forever grateful.
	\\ \indent Finally thanks to my brother and parents. Coming from a family of
	computer scientists certainly rubs off and because of all their love and support I have
	been able to develop my own passions.}

\ToCisShort	% use this only for 1-page Table of Contents

\LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

\LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%       BEGIN DOCUMENT...         %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\input macros.tex
\input chapter1.tex
\input chapter2.tex

%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{plain}	% or "siam", or "alpha", etc.
\nocite{*}		% list all refs in database, cited or not
\bibliography{refs}		% Bib database in "refs.bib"


%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
\appendix
\input appendixA.tex
\input appendixB.tex


\end{document}

